---
title: "Getting started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  warning = F, 
  message = F, 
  fig.align = "center"
)
```

## Installation

Install the package from [r-universe](http://tbep-tech.r-universe.dev/ui/#builds) as follows. The source code is available on [GitHub](https://github.com/tbep-tech/slrcsap).

```{r eval = F}
# Install slrcsap in R:
install.packages('slrcsap', repos = c('https://tbep-tech.r-universe.dev', 'https://cloud.r-project.org'))
```

Load the package in an R session after installation:

```{r}
library(slrcsap)
```

## Usage

The package includes two core functions to download data and two core functions to plot data.  The two download functions are `get_sealevel()` and `get_scenario()` to download tidal gauge data and sea level rise scenario data, respectively.  The corresponding plot functions to view data retrieved from the download functions are `plot_sealevel()` and `plot_scenario()`. The content below demonstrates how to use each function, including the optional arguments that can be used to customize the output.

Tidal gauge data is downloaded from the NOAA Tides and Currents website.  The data is available for all NOAA tide gauges, but the package is currently set up to download data for the St. Petersburg, FL gauge (NOAA ID 8726520).  The data is downloaded as a CSV file and then read into R.  The data includes monthly mean sea level (MSL) values from 1947 to the present, including a seasonal correction.

```{r}
# Download tidal gauge data
dat <- get_sealevel()
head(dat)
```

